---
title: "MATH400FinalProject"
author: "Melissa Kimble"
date: "April 28, 2019"
output: html_document
---

```{r setup}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)

LibraryList<-c("ggplot2","mgcv","tidyverse","leaps")
for (TheLibrary in LibraryList)
{
  if(TheLibrary %in% rownames(installed.packages()) == FALSE) install.packages(TheLibrary)
}
library(ggplot2)
library(mgcv)
library(tidyverse)
library(leaps)

FileNameDate<-as.character(format(Sys.Date(), "%Y%m%d"))
TodayDate<-as.character(format(Sys.Date(), "%m-%d-%Y"))

InputDataFolderPath<-"D:/Dropbox/00_Dissertation/01_Data/02_Working/WaterQuality/"
OutputCSVFolderPath<-"D:/Dropbox/00_Dissertation/01_Data/02_Working/WaterQuality"
OutputDataFolderPathPNG<-paste("D:/Dropbox/00_Dissertation/01_Data/04_Figures/Aquaculture/State_DMR/From_Geoff/",FileNameDate,"/",sep="")

TheDirs=c(OutputCSVFolderPath,OutputDataFolderPathPNG)

## checks if the the directories exist, if they don't it creates them.
for (TheDir in TheDirs){
  DirectoryExists<-dir.exists(file.path(TheDir))
  if (DirectoryExists==FALSE) dir.create(file.path(TheDir), recursive = TRUE)
}

## WQ and WL were cleaned and updated (appended new dates) in a separate R script
all_DBFMerge<-read.csv(paste(InputDataFolderPath,"all_DMRBuoyRast_UTM.csv",sep=""))
colnames(all_DBFMerge)[colnames(all_DBFMerge)=="LATITUDE_D"] <- "LATITUDE_DECIMAL"
colnames(all_DBFMerge)[colnames(all_DBFMerge)=="LONGITUDE_"] <- "LONGITUDE_DECIMAL"
all_DBFMerge<-all_DBFMerge[,c("Id","Station","LATITUDE_DECIMAL","LONGITUDE_DECIMAL","Date","YrMo","Easting","Northing","TempC","Sal",
                              "WGOM_m","Type","DMRBuoyID","SMAFishId","SatTempC","StreamDist","Zone","Depth_m")]
all_DBFMerge$DateTime<-as.POSIXlt(all_DBFMerge$Date, tz="GMT", format="%m/%d/%Y %H:%M:%S")
all_DBFMerge$Date2<- strftime(all_DBFMerge$DateTime,"%Y-%m-%d")
all_DBFMerge$Date<-all_DBFMerge$Date2
ColDrop <- c("Date2","DateTime")
all_DBFMerge<-all_DBFMerge[ , !(names(all_DBFMerge) %in% ColDrop)]

all_DBFMerge$Month<-format(as.Date(all_DBFMerge$Date, "%Y-%m-%d"),"%b")
all_DBFMerge$Year<-format(as.Date(all_DBFMerge$Date, "%Y-%m-%d"),"%Y")

Months<-c("Jan","Feb","Mar","Apr","May","Jun","Jul","Aug","Sep","Oct","Nov","Dec")
all_DBFMerge$Month<-factor(all_DBFMerge$Month, levels = Months)

all_DBFMerge_YrMoSub<-all_DBFMerge
#all_DBFMerge_YrMoSub<-all_DBFMerge[(all_DBFMerge$YrMo2>="2014-05" & all_DBFMerge$YrMo2<="2014-11"),]

all_DBFMerge_YrMoSub<-all_DBFMerge_YrMoSub[order(all_DBFMerge_YrMoSub$Date),]
rownames(all_DBFMerge_YrMoSub) <- seq(length=nrow(all_DBFMerge_YrMoSub))
head(all_DBFMerge_YrMoSub)

```

## Background Information
Every data source has characteristic uncertainties. Spatially accurate water temperature estimates aid in identifying suitable aquaculture sites but are challenging to obtain due to limited data availability and because available sources on their own tend to be biased. I am investigating methods for combining data that focus on reducing bias and uncertainty in individual sources. The target products are monthly coastal water temperature and salinity maps that provide a baseline of temperature values for each grid cell. 

The Department of Marine Resources collects randomly stratified samples of sea surface temperature and salinity every month since 1990. These samples, however, are only taken along the coast, so values for temperature and salinity for the remainder of the estuary have to be obtained from other sources. The New Meadows (zone WL) and Damarscotta River (zone WQ) have two to three LOBO Buoys each that have been deployed from 2015 - 2018. They collect several oceanographic variables which include sea surface temperature and salinity. The buoys are generally in deeper water near the center of their respective estuaries. The UMaine Satellite Oceanography group routinely converts Landsat-8 into sea surface temperature, which range from 2013 to 2016. The temporal and spatial resolution and the coverage of these datasets vary. Since reflectance obtained from satellite imagery is essentially an average of an area on the ground (for LSAT8, 100-meter), nearshore values are biased because of land influence. Additionally, the DMR samples are susceptible to fresh water inputs and land based run-off. Other datasets used in this study were the Western Gulf of Maine 8-meter bathymetry and euclidean distance from streams derived from the National Hydrography Dataset. Due to the diversity of data types (rasters, vector, and tabular), an 100-meter grid was overlaid and values were averaged within each grid. The sampling points and buoy locations did not overlap grid cells, but the streams, bathymetry, and LSAT8 sea surface temperature values were averaged within any shared grids.  

Initially, satellite derived sea surface temperature values were compared to sampled sites and on average, water quality monitoring samples were taken within approximately 10 days of the satellite scene’s collection date. The discrepancy in collection dates were as distant as 29 days, to as close as the same day. About 7% of monitoring sites overlapped with the satellite scenes, and of the 35 evaluated scenes, 18 were within the extent of the monitoring sites for years 2013 to 2016. High sea surface temperature values in warmer months tended to correspond with lower monitoring values, and vice versa for colder months when separated by month and year and by month. The average difference in temperature was 6C. The satellite derived sea surface temperature has nearly full coverage of each estuary, so the goal of the comparison was to see if the two datasets were significantly different. For most months they were either significantly different or their distributions were not similar. This may have been due to two main discrepancies in the datasets, however. Several sampling points did not overlap with the satellite data, so the nearest set of pixel values were sampled and averaged. Lastly, only 8 satellite scenes of 35 had days with matching sampling and buoy dates. The overlap primarily consisted of buoy locations, which correlated well (<.90) with satellite derived sea surface temperature. This correlation, however, is expected due to buoys located in deeper waters with less land-based influence. For this project the focus was primarily on the collected salinity and temperature variables from sampling sites and buoy locations.

## Exploratory Data Analysis

Water Quality Zones WQ and WL cover different regions, so the below plots were to check if the distributions of Salinity and Temperature were notably different. Zone will be included as a separate variable, because salinity in water quality zone WL has short tails while WQ has a left skew. Temperature for both regions were not notably different.
There is high correlation between satellite temperature and buoy locations, which is to be expected because the buoy locations are closer to the center of the estuary, less susceptible to fresh water or land-based inputs.

```{r}
range(all_DBFMerge$Year[all_DBFMerge$Zone=="WQ"]);range(all_DBFMerge$Year[all_DBFMerge$Zone=="WL"])
range(all_DBFMerge$TempC[all_DBFMerge$Zone=="WQ"]);range(all_DBFMerge$TempC[all_DBFMerge$Zone=="WL"])
range(all_DBFMerge$Sal[all_DBFMerge$Zone=="WQ"]);range(all_DBFMerge$Sal[all_DBFMerge$Zone=="WL"])

corrTest<-na.omit(all_DBFMerge)
cor(as.numeric(corrTest$TempC),as.numeric(corrTest$SatTempC))

par(mfrow =c(1,2))
# by Zone - WQ or WL
ggplot(all_DBFMerge_YrMoSub,aes(x=as.numeric(Sal))) + 
  ylab("Density") +
  xlab("Salinity (PPT)") +
  geom_histogram(aes(y=..density..), colour="black", fill="white")+
  geom_density(alpha=.2, fill="#FF6666") +
  facet_grid(Zone ~ .) 
# by Type - Sample or Buoy location
ggplot(all_DBFMerge_YrMoSub,aes(x=as.numeric(Sal))) + 
  ylab("Density") +
  xlab("Salinity (PPT)") +
  geom_histogram(aes(y=..density..), colour="black", fill="white")+
  geom_density(alpha=.2, fill="#FF6666") +
  facet_grid(Type ~ .) 
```

```{r}
par(mfrow =c(1,2))

# by Zone - WQ or WL
ggplot(all_DBFMerge_YrMoSub,aes(x=as.numeric(TempC))) + 
  ylab("Density") +
  xlab("Temperature (C)") +
  geom_histogram(aes(y=..density..), colour="black", fill="white")+
  geom_density(alpha=.2, fill="#FF6666") +
  facet_grid(Zone ~ .)
# by Type - Sample or Buoy location
ggplot(all_DBFMerge_YrMoSub,aes(x=as.numeric(TempC))) + 
  ylab("Density") +
  xlab("Temperature (C)") +
  geom_histogram(aes(y=..density..), colour="black", fill="white")+
  geom_density(alpha=.2, fill="#FF6666") +
  facet_grid(Type ~ .)
```

```{r}
par(mfrow =c(1,2))

# temperature boxplots by zone
ggplot(all_DBFMerge_YrMoSub, aes(x=Zone, y=as.numeric(TempC), fill=Zone)) + 
  geom_boxplot() +
  ylab("Temperature (C)") +
  xlab("Zone") +
  facet_grid(Type ~ .)

# salinity boxplots by zone
ggplot(all_DBFMerge_YrMoSub, aes(x=Zone, y=as.numeric(Sal), fill=Zone)) + 
  geom_boxplot() +
  ylab("Salinity (PPT)") +
  xlab("Zone") +
  facet_grid(Type ~ .)
```

WL and WQ do not not have extreme differences in the relationship between Salinity ~ Temperature. Season, however, will need to be treated separately as a variable (by month) within the shellfish water quality monitoring dataset.


```{r}
par(mfrow =c(2,2))
# salinity ~ temperature by type & zone
ggplot(data = all_DBFMerge_YrMoSub, aes(x = as.numeric(TempC), y = as.numeric(Sal), fill=Zone, colour=Zone)) +  
  geom_point() + 
  geom_smooth() + 
  xlab("Temperature (C)") + 
  scale_x_continuous(breaks=seq(0,35,5)) +
  scale_y_continuous(limits=c(0,35),breaks=seq(0,35,5)) +
  ylab("Salinity (PPT)") +
  ggtitle("Salinity ~ Temperature") +
  theme(plot.title = element_text(hjust=0.5)) +
  facet_grid(Type ~ .)

# salinity ~ temperature by zone & month
ggplot(data = all_DBFMerge_YrMoSub, aes(x = as.numeric(TempC), y = as.numeric(Sal), fill=Month, colour=Month)) +  
  geom_point() + 
  geom_smooth() + 
  xlab("Temperature (C)") + 
  scale_x_continuous(breaks=seq(0,35,5)) +
  scale_y_continuous(limits=c(0,35),breaks=seq(0,35,5)) +
  ylab("Salinity (PPT)") +
  ggtitle("Salinity ~ Temperature") +
  theme(plot.title = element_text(hjust=0.5)) +
  facet_grid(Type ~ .)

```



```{r}
set.seed(1)
all_DBFMerge_modsub<-all_DBFMerge[,c("Month","Easting","Northing","TempC","Sal","WGOM_m","Type","StreamDist","Zone","Depth_m")]
rp <- modelr::resample_partition(all_DBFMerge_modsub, c(train = 0.7, test = 0.3))
training_set <- as_tibble(rp$train)
testing_set <- as_tibble(rp$test)

regfit_fwd=regsubsets(Sal~.,data=training_set, nvmax=20, method="forward")
regfit_summ<-summary(regfit_fwd)

MinRSS<-which.min(regfit_summ$rss)
MinRSS
coef(regfit_fwd,MinRSS) 

MaxAdjr2<-which.max(regfit_summ$adjr2)
MaxAdjr2
coef(regfit_fwd,MaxAdjr2)

MinCP<-which.min(regfit_summ$cp)
MinCP
coef(regfit_fwd,MinCP) 

MinBIC<-which.min(regfit_summ$bic)
MinBIC
coef(regfit_fwd,MinBIC) 

par(mfrow =c(2,2))
plot(regfit_summ$rss, xlab="Number of Variables", ylab="RSS", type="l")
points(MinRSS, regfit_summ$rss[MinRSS], col =" red",cex =2, pch =20)
plot(regfit_summ$adjr2, xlab="Number of Variables", ylab="Adjusted RSq",type="l")
points(MaxAdjr2, regfit_summ$adjr2[MaxAdjr2], col ="red", cex=2, pch =20)
plot(regfit_summ$cp ,xlab ="Number of Variables",ylab="Cp",type="l")
points(MinCP, regfit_summ$cp[MinCP], col="red",cex=2, pch =20)
plot(regfit_summ$bic, xlab="Number of Variables", ylab="BIC", type="l")
points(MinBIC, regfit_summ$bic[MinBIC], col =" red",cex =2, pch =20)

## Based on the evaluation plots, model performance stops making gains at 10 coefficients.
coef(regfit_fwd,10) # MonthMar, MonthApr, MonthMay, MonthJun, Easting, Northing, TempC, TypeSample, StreamDist, ZoneWQ 
training_set<-training_set[,c("Month","Easting","Northing","TempC","Sal","Type","StreamDist","Zone")]
testing_set<-testing_set[,c("Month","Easting","Northing","TempC","Sal","Type","StreamDist","Zone")]

```

It was anticipated that R-sq.(adj) =  0.349   Deviance explained = 35.1%

```{r}
## Base model, no splines.
fit_gam <- mgcv::gam(Sal ~ Month+Easting+Northing+TempC+Type+StreamDist+Zone, data = training_set)
par(mfrow =c(3,3))
mgcv::plot.gam(fit_gam, all.terms=TRUE, se=TRUE)

## TempC 
fit_gam1 <- mgcv::gam(Sal ~ Month+Easting+Northing+s(TempC,k=-1,bs="cs")+Type+StreamDist+Zone, data = training_set)
anova(fit_gam,fit_gam1,test="F")
## spline significant on TempC

## StreamDist 
fit_gam2<- mgcv::gam(Sal ~ Month+Easting+Northing+s(TempC,k=-1,bs="cs")+Type+s(StreamDist,k=-1,bs="cs")+Zone, data = training_set)
anova(fit_gam1,fit_gam2,test="F")
## spline significant on StreamDist

## Easting 
fit_gam3<- mgcv::gam(Sal ~ Month+s(Easting,k=-1,bs="cs")+Northing+s(TempC,k=-1,bs="cs")+Type+s(StreamDist,k=-1,bs="cs")+Zone, data = training_set)
anova(fit_gam2,fit_gam3,test="F")
## spline significant on Easting

## Northing 
fit_gam4<- mgcv::gam(Sal ~ Month+s(Easting,k=-1,bs="cs")+s(Northing,k=-1,bs="cs")+s(TempC,k=-1,bs="cs")+Type+s(StreamDist,k=-1,bs="cs")+Zone, data = training_set)
anova(fit_gam3,fit_gam4,test="F")
## spline significant on Northing

```
```{r}
## Final model
par(mfrow =c(3,3))
mgcv::plot.gam(fit_gam4, all.terms=TRUE, se=TRUE)

summary(fit_gam4)

par(mfrow = c(2,2))
gam.check(fit_gam4)

```
## Results & Discussion
When salinity is not available, temperature is often used as a substitute because the relationship between the two are well understood in relatively stable conditions. This substitution becomes problematic in estuarine environments, where there are frequent inputs from fresh water and run-off from nearshore environments. Based on this project, it would seem as though publications that substitute temperature for salinity should not do so in estuarine environments. I was not able to get fresh water inputs available in time for the project due to the complexity of all of the datasets that needed to be synthesized, but included those listed in the introductory of this document. I also want to explore bayesian approaches and need to ensure any testing and training datasets are independent. Spatial datasets tend to be autocorrelated, which would bias diagnostics and model fit. Within this project I was not able to complete a sampling scheme for ensuring sampled datasets are not autocorrelated other than running a Moran's I on every sampled dataset, which is not efficient or feasible. Right now I am working with using distance from points as a way to sample points, but this works with pairs and will need to be tweaked for the entire dataset to not bottleneck the process. I also explored quadrants, but the issue within th equadrants still remains and simple random stratified sampling did not seem to work as well as needed.

Overall, there was extremely poor GAM performance (<.30 adj r^2) and an obvious skew to model residuals. I do not think this was because of the model chosen, but because temperature is simply not a good explanatory variable for salinity in estuarine environments. I do think freshwater quantity and velocity and currents will improve predictions, however. 

```{r}
test_sub <- testing_set %>% select(Month, Easting, Northing, TempC, Sal, Type, StreamDist, Zone)
# Report the test error obtained
# https://drsimonj.svbtle.com/visualising-residuals
test_sub$predicted<-predict(fit_gam4, testing_set)
test_sub$residuals<-test_sub$predicted-testing_set$Sal
hist(test_sub$residuals)

gam_test_error<-mean((test_sub$residuals)^2)
gam_test_error # 15.97974

test_sub %>% 
  gather(key = "var", value = "value", -Sal, -predicted, -residuals, -Month, -Type, -Zone) %>%  # Get data into shape
  ggplot(aes(x = value, y = Sal)) +  # Note use of `x` here and next line
  geom_segment(aes(xend = value, yend = predicted), alpha = .2) +
  geom_point(aes(color = residuals)) +
  scale_color_gradient2(low = "blue", mid = "white", high = "red") +
  guides(color = FALSE) +
  geom_point(aes(y = predicted), shape = 1) +
  facet_grid(~ var, scales = "free") +  # Split panels here by `var`
  theme_bw()

# https://www.r-graph-gallery.com/265-grouped-boxplot-with-ggplot2/
group_melt<-test_sub %>% 
  gather(key = "group", value = "groupVal", -Month, -Easting, -Northing, -TempC, -Type, -StreamDist, -Zone, -residuals) 
var_melt<-group_melt %>%   
  gather(key = "var", value = "value", -group, -groupVal, -residuals)
ggplot(var_melt,aes(x=var, y=groupVal, fill=group)) + 
    geom_boxplot() +
    facet_wrap(~var, scale="free")

```