---
output:
  html_document: default
  pdf_document: default
---
```{r}
library(MASS)
library(tidyverse)
library(ISLR)
head(Smarket)
```

# Logistic Regression

```{r}
glm_fit <- glm(Direction ~ Lag1 + Lag2,
    data = Smarket,
    family = binomial
)
glm_fit
```

```{r}
library(modelr)
new_data <- read_csv("Lag1, Lag2
                      0.5, 0.3
                      0.4, 0.3")
new_data %>% add_predictions(glm_fit) %>% mutate(prob = exp(pred)/ (1 + exp(pred)))
```

# ROC curve

```{r}
Smarket2 <- Smarket %>% 
    add_predictions(glm_fit) %>% 
    mutate(prob = exp(pred)/ (1 + exp(pred)), EstDir = ifelse(prob > 0.5, "Up", "Down"))
Smarket2 %>% count(Direction, EstDir) %>% spread(Direction, n)
```

```{r}
library(tidymodels)
autoplot(roc_curve(Smarket2, Direction, prob))
roc_auc(Smarket2, Direction, prob)
```

# Multinomial logistic


variable name |  type |  about the variable 
--------------|--------|-------------------------------------------------------
id          |    scale|  student id                  
female      |  nominal|  (0/1)        
race        |  nominal|  ethnicity (1=hispanic 2=asian 3=african-amer 4=white)        
ses         |  ordinal|  socio economic status (1=low 2=middle 3=high)           
schtyp      |  nominal|  type of school (1=public 2=private)
prog        |  nominal|  type of program (1=general 2=academic 3=vocational)
read        |    scale|  standardized reading score
write       |    scale|  standardized writing score
math        |    scale|  standardized math score
science     |    scale|  standardized science score
socst       |    scale|  standardized social studies score
hon         |  nominal|  honors english (0/1)

```{r}
ml <- read_csv("hsb2.csv")
```

```{r}
hsb2 <- ml %>% mutate(
    prog = recode_factor(prog, `1` = "general", `2` = "academic", `3` = "vocational"),
    ses = recode_factor(ses, `1` = "low", `2` = "middle", `3` = "high"))
```

```{r}
hsb2 %>% count(prog, ses)  %>% spread(prog, n)
```

```{r}
hsb2 %>% group_by(prog) %>% summarize(mwrite = mean(write))
```

```{r}
library(nnet)
multi_fit <- multinom(prog ~ ses + write, data = hsb2)
```

```{r}
new_data <- tibble(ses = "middle", write = 56)
predict(multi_fit, new_data, type = "probs")


# tidyverse
new_data %>% add_predictions(multi_fit)
```

```{r}
predict(multi_fit, new_data)

# a few month later
# new_data %>% add_predictions(multi_fit, type = "probs")
```

# LDA
```{r}
lda_fit <- lda(Direction ~ Lag1 + Lag2, data = Smarket)
```

```{r}
Smarket3 <- Smarket %>% mutate(EstDir = predict(lda_fit, newdata = Smarket)$class)
Smarket3 %>% count(Direction, EstDir) %>% spread(Direction, n)
```


```{r}
new_data <- read_csv("Lag1, Lag2
                      0.5, 0.3
                      0.4, 0.3")

predict(lda_fit, new_data)

# modelr magic doesn't work now
# new_data %>% add_predictions(lda_fit)
```

## An exmple with more than one class

```{r}
head(iris)
```

```{r}
iris_fit <- lda(Species ~ .,  data = iris)
```

```{r}
new_data <-  tibble(Sepal.Length = 5.906, Sepal.Width = 2.77, Petal.Length = 3, Petal.Width = 0.246)
predict(iris_fit, new_data)$posterior
```

```{r}
iris2 <- iris %>% bind_cols(as_tibble(predict(iris_fit, newdata = iris)$x))
ggplot(iris2) + geom_point(aes(LD1, LD2, color = Species))
```

# Reduced rank LDA

```{r}
# only use LD1
predict(iris_fit, new_data, dimen = 1)
```

# Explaining kNN

```{r}
library(mvtnorm)
g1 <- rmvnorm(30, mean=c(-0.5,0))
g2 <- rmvnorm(20, mean=c(1,-1))
g3 <- rmvnorm(20, mean=c(1,1))
x <- rbind(g1, g2, g3)
colnames(x) <- c("x1", "x2")
x <- as_tibble(x)
y <- rep(c("a", "b", "c"), c(30, 20, 20))
knn_example <- bind_cols(x, y = y)
```

```{r}
ggplot(knn_example) + geom_point(aes(x1, x2, color = y))
```

```{r}
k = 7
point = c(-1, 0)
knn_neighor <- knn_example %>% 
  mutate(dist = sqrt((x1 - point[1])^2 + (x2 - point[2])^2)) %>%
  filter(row_number(dist) <= k)
ggplot(knn_example) +
  geom_point(data = knn_neighor, aes(x1, x2), alpha = 0.5, size = 5) +
  geom_segment(data = knn_neighor, aes(x = x1, y = x2, xend = point[1], yend = point[2], color = y)) +
  geom_point(aes(x1, x2, color = y)) +
  annotate("point", x = point[1], y = point[2])
```


```{r}
library(class)
new_data <- read_csv("x1, x2
                      0.2, 1
                      0.6, -1")
new_data %>% mutate(est_class = knn(knn_example %>% select(x1, x2), new_data, knn_example %>% pull(y), k = 5))
```


